Identifying and correcting multiple sources of misspecification\\in GWAS summary statistics for prediction\\---~OR~---\\Reducing misspecifications and improving tagging\\for better polygenic scores

### 2021-08-17 paper/paper-misspec.tex
desc larger sets

[TODO: MAYBE THE RIGHT PLACE TO PRESENT THE TWO NEW SETS]

>>>>

We also design two larger sets of imputed variants to compare against using only HapMap3 variants for prediction. 
Following \cite{prive2021high}, we first restrict to UKBB variants with MAF > 0.01 and INFO > 0.3.
We then compile frequencies and imputation INFO scores from other datasets, iPSYCH and summary statistics for breast cancer, prostate cancer, coronary artery disease, type-1 diabetes and depression \cite[]{bybjerg2020ipsych2015,michailidou2017association,schumacher2018association,nikpay2015comprehensive,censin2017childhood,wray2018genome}. 
We restrict to variants with a mean INFO > 0.3 in these other datasets, and also compute the median frequency per variant.
To exclude potential mismappings in the genotyped data \cite[]{kunert2020allele} that might have propagated to the imputed data, we compare median frequencies in the external data to the ones in the UK Biobank.
As we expect these potential errors to be localized around errors in the genotype data, we apply a moving-average smoothing on the frequency differences to increase power to detect these errors and also reduce false positives.
We define the threshold on these smoothed differences based on visual inspection of their histogram (0.03).
This results in an initial set of 9,394,361 variants.

From this large set of variants, we define two sets.
One is based on clumping, using a threshold $r^2 = 0.9$ over a radius of 100 Kbp and prioritizing HapMap3 variants and larger INFO scores. This results in a set ``clump'' of 2,465,478 variants, among which there are 554,655 of the initial HapMap3 variants.
For the second set, we aim at maximizing the tagging of the initial 9,394,361 variants, i.e. $\sum_{i \in \text{all}} \max_{j \in \text{set}}{|R_{i,j}|}$ (inspired from the alternative sensitivity of \cite{agier2016systematic}). 
We design a greedy algorithm that first select all HapMap3 variants, then add one variant at a time, the one that maximizes the addition to this sum, until no variant can add more than 0.2.
This results in a set ``maxtag'' of 2,029,086 variants.
